Lattice Boltzmann study of mass transfer for Bretherton/Taylor bubble train flow

### 2012-04-30 Text/introduction.tex
The iteration from my side is finished.

\begin{document}
\maketitle
\begin{abstract}
This work presents a thorough procedure of how to determine the volumetric mass transfer
coefficient
in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train flow. In this work we address the situation when the hydrodynamic pattern changes from having a vortex in slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}. For the latter case the bubble shape is not symmetric and cannot be approximated through flat surfaces and circumferences as it is done in literature \cite{vanbaten-circular,kreutzer-overview}. Moreover, with a vortex in the slug the tracer is well mixed. Thus, it is common to  use periodic boundary conditions  and inlet/outlet averaged concentration as the characteristic concentration, Eq. \ref{eq:main:definition}.  The later is not valid for flows where tracer is not well mixed. This work through examination of different boundary conditions and different characteristic concentration definitions establishes the procedure to determine the volumetric mass transfer coefficient for flows with the capillary number $0.1<Ca<1.0$. With simulations of a few unit cells the connection between time (used in simulations) and space (used in experiments) domains was established. The following boundary conditions were examined: periodic boundary conditions, open boundaries, and a few unit cells open boundary
simulations. It was shown that the time dependent average concentration to be a characteristic
concentration produce robust results, though a bit overpredicted. The inlet/outlet flux averaged
concentration \cite{vanbaten-circular} to be a characteristic concentration produce unreliable results in the range of capillary numbers specified. In comparison with works \cite{vanbaten-circular,kreutzer-overview} where only one unit cell was simulated, a few
unit cells open boundaries conditions resemble physical situation most closely and are harder to conduct
but provide the volumetric mass transfer coefficient based only on the spatial information for the
capillary number $Ca>0.7$. 
%As well, a few unit cells simulations can provide more information about
%mixing in the liquid slug in comparison with the periodic boundary conditions. 
We show
that all presented in literature strategies are extreme limits of one equation,
Eq. \ref{main:main:main}. Finally, simulation results for different Peclet numbers
are compared with analytical predictions presented by \citet{vanbaten-circular} and shows a good
agreement. The work is of use to people performing mass transfer numerical simulations for bubble
train flow in microchannels and/or within the
lattice Boltzmann context.
\end{abstract}

\section{Introduction}
\label{intro}
The monolith reactors during last years are getting more attention as a promising alternative to slurry
reactors and trickle bed reactors \cite{kreutzer-overview,bercic-mass}. Reactors usually operate in Bretherton-Taylor regime \cite{bretherton,taylor} which describes a flow through
a liquid medium of equally sized, long bubbles, see
Fig. \ref{fig:benchmark:hydro}. This flow regime
is characterized by surface tension dominance over inertia and viscous effects, and by
comparatively small gas flow velocities \cite{yue-mass}. Due to surface tension dominance bubble train flow exhibit advantageous properties which can not be achieved
in macroscopic counterparts: liquid thin films \cite{bretherton} between bubbles and walls remarkably enhance mass
transfer from a gas and walls to liquid; the plug flow regime occurring in monolith reactors allows to perform chemical reactions in slugs only \cite{kreutzer-overview}. Moreover, the low slip velocity between gas and liquid is utilized in the experiments to measure liquid velocity \cite{taylor}:  bubbles travelling with approximately the same velocity as liquid can be captured with a camera. Overall, nowadays one can find a large number of applications
of the Bretherton-Taylor bubble train flow: continuous flow analyzers to measure liquid velocity, lung openings, chemical reactions
as hydrogenation of nitroaromatics, 2-ethyl-hexenal, Fischer-Tropsch synthesis, etc. Extensive reviews of \citet{kreutzer-overview,
gupta-review,yue-mass} cover a significant number of applications.

This work is focused on the gas to liquid mass transfer 
study for the Bretherton/Taylor flow. The good understanding of mass transfer and how it depends on the
parameters as the capillary number, the Reynolds number, slug and bubble lengths allows to properly
manufacture a microchannel with necessary properties to ensure that chemical
reactions are performed in the best manner. The mass transfer coefficient characterizing mass transfer is defined as the flux from the surface divided on the concentration difference between the concentration imposed on the body and the characteristic concentration in the domain. The characteristic concentration in the domain depends largely on underlying hydrodynamics fields.  For example, experimental studies \cite{yue-mass,bercic-mass} show a complex dependency of the mass transfer coefficient on flow parameters: bubble and slug lengths, bubble velocity, which are in turn depend on the capillary number $Ca$ and the Reynolds number $Re$. 
\citet{yue-mass} established an experimental correlation for the volumetric mass transfer coefficient for a bubble train as
a function of the diffusion coefficient, slug and bubble lengths, and bubble velocity: 
\begin{equation}
\vol =\frac{2}{d_h} \Bigl(\frac{D
\ububble}{\lbubble+\lslug}\Bigr)^{0.5}
\Bigl(\frac{\lbubble}{\lbubble+\lslug}\Bigr)^{0.3},
\end{equation}
where $\vol$ is the volumetric mass transfer coefficient, $d_h$ is the hydrodynamic capillary
diameter, $\lbubble$ is the bubble length, $\lslug$ is the slug distance (between bubbles),
$\ububble$ is the bubble velocity, $D$ is the diffusion coefficient. 
%This correlation is within $10\%$ accuracy for experimental
%results for the microchannels with rectangular/square crosssections for the following range of
%parameters: bubble velocity $0.4\,\mathrm{m/s}<\ububble<2\,\mathrm{m/s}$, bubble length
%$1.4<\frac{\lbubble}{d_h}<6.3$, slug length $1<\frac{\lslug}{d_h}<3.2$,
%hydrodynamic diameter $d_h = 400 \mathrm{\mu m}$.

Thus, the understanding of mass transfer for the bubble train flow is not possible without deep understanding of hydrodynamic patterns. There are a number of studies available for the hydrodynamic study of the bubble train flow: experimental \cite{kreutzer-pressure-drop,cerro-space,cerro-bubble-train} and numerical \cite{wang-non-circular,kuzmin-binary3d,giavedoni-numerical,heil-threedim}. For the flow of long bubbles between parallel plates as the study case it is indicated that  there exists a vortex in the liquid slug for $Ca<0.7$ , see Fig. \ref{fig:streamlines:tweaked:velocity}. However, for $Ca>0.7$ the vortex in the liquid slug does not exist. As well the bubble shape for low capillary numbers ($Ca<0.1$ \cite{cerro-bubble-train}) is symmetric. Thus, in the case of flow between plates the bubble shape can be represented as two hemicircles and two planar interfaces. This is heavily utilized for mass transfer analytical estimations.

As mentioned before the mass transfer coefficient is defined as the mass flux from certain area, Eq. \ref{eq:main:definition}. Thus, analytical estimations \cite{kreutzer-overview,irandoust} are  based on decomposition of the bubble shape. The mass transfer coefficient is calculated through two separate contributions from two films and two hemicircles. For both contributions the Higbie penetration theory \cite{higbie} is utilized, which states that the mass transfer coefficient  for simple geometry flow depends on the average time of liquid packet interaction with a geometrical feature. It can be calculated as $\sqrt{\frac{\pi D}{t_{\mathrm{char}}}}$, where $t_{\mathrm{char}}$ is the interaction characteristic time, $D$ is the diffusion coefficient. As the example for the Higbie penetration theory, the mass  transfer coefficient for the flow of bubbles between parallel plates  is calculated as (similarly to work of \citet{vanbaten-circular}):
\beq
k_L=2 \sqrt{\frac{\pi D}{t_{\mathrm{film}}}}+2 \sqrt{\frac{\pi D}{t_{\mathrm{circle}}}},
\feq
where $t_{\mathrm{film}}=\frac{\lfilm}{\ububble}$ states for the interaction time of liquid traveling through planar part of the bubble, $t_{\mathrm{circle}}=\frac{\pi R_{\mathrm{circle}}}{\ububble}$ is the time which liquid in the slug travels the distance of half circumference.

@@

However, available numerical studies of mass transfer \cite{kreutzer-overview,vanbaten-circular} are lacking the simulation of  bubble shapes. The usual simulation of the mass transfer  is performed as follows: 
\begin{description}
\item[I] The bubble shape is calculated either through analytical correlations \cite{bretherton} or experimental correlations \cite{cerro-bubble-train} without directly resolving bubbles shapes through multiphase simulations. 
\item[II] Hydrodynamics fields are then obtained by performing simulations of one-component flow around the droplet by imposing the bubble velocity on the walls. Thus, the simulations are performed in the reference frame moving with the bubble. The stress free condition is imposed at the bubble surface. 
\item[III]  The mass transfer simulations are  performed in the reference frame moving with the bubble. The saturation concentration is imposed at the bubble surface. Only one unit cell with one bubble in it is used for simulations. Periodic boundary conditions are utilized to determine the volumetric mass transfer coefficient, which is calculated through the following equation \cite{vanbaten-circular}:
\begin{equation}
\label{main:simulation:equation}
\vol=\frac{\mathrm{\overline{Flux}}}{\cbubble-\langle\coutlet\rangle} \frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\
cell\ volume}},
\end{equation}
where $\langle\coutlet(t)\rangle=\int{C \uoutlet \mathrm{d}A}/\int{\uoutlet\mathrm{d}A}$ is the averaged in space outlet concentration as the function of time used as the characteristic concentration in the definition, Eq. \ref{eq:main:definition}. 
The averaged in time concentration flux ($\mathrm{\overline{Flux}}$) is calculated as the difference between the overall
average concentration in the whole domain ($\langle\coverall\rangle=\int_{V} C \mathrm{d}V /V$)
at time
$t_1$ and at time $t_2$ divided on the time difference $t_2-t_1$. The agreement between numerical
simulations \cite{vanbaten-circular} and correlations of \citet{bercic-mass} was good. 
\end{description}
There is a certain critics towards the presented numerical approaches \cite{vanbaten-circular,kreutzer-overview}. They mainly originate from the bubble shape approximation. It is taken symmetric, i.e. consisting from hemispheres and cylinder film for the case of flow in circular capillaries. This is valid  for small capillary numbers $Ca<0.1$. As it was discussed, for such capillary numbers in the reference frame moving with the bubble there is a vortex in slug. Thus, the tracer is well mixed in slug.The choice of the characteristic concentration needed for the mass transfer coefficient, Eq. \ref{eq:main:definition}, in this case  is obvious. With small differences in results it can be either averaged in the liquid slug concentration or the outlet space averaged concentration as used in formulation of \citet{vanbaten-circular}. Another critics is towards periodic boundary conditions to calculate the mass transfer coefficient. While it is clear to use periodic boundary conditions for the calculation of hydrodynamics fields for  long bubbles motion in the microchannel, it is not the case for mass transfer simulations. Experimental correlations \cite{bercic-mass} show that the concentration needed for the mass transfer calculations along the streamwise direction changes as the exponential function. However, mass transfer simulations are made only for one unit cell using the periodic boundary conditions with the same concentration at the inlet and at the outlet. The question how  one unit cell simulation corresponds to experimental measurements arises where concentration difference is measured at the distances of at least of a few unit cells difference \cite{bercic-mass}. In other words one needs to understand how the discrete one unit cell simulation corresponds to the continuous picture in experiments where one does not distinguish bubbles but takes measurements of concentration at different locations.

@@

\item
The approach of \citet{vanbaten-circular} is examined. Periodic boundary conditions were used. The work \cite{vanbaten-circular}
calculates the mass transfer coefficient as the gain of the mass in the system
divided by the concentration difference
multiplied on the surface area:
\beq
\vol=\frac{\dot{m}}{P \Delta C}\frac{P}{V}=\frac{\dot{m}}{V (\cstar-\langle C(t) \rangle)},
\feq
where mass difference in the domain can be calculated as:
\beq
\dot{m}=\frac{m_2-m_1}{t_2-t_1}=\frac{\int_{liq}{C(\bm{x},t_2)\mathrm{d}V}-\int_{liq}{
C(\bm{x},t_1)\mathrm { d } \bm{x}} } {
t_2-t_1 }.
\feq
In the approach of \citeauthor{vanbaten-circular} the inlet and outlet concentrations were taken as the
characteristic concentration $\langle C(t) \rangle $.

@@

\section{Validation}
For analytical expressions the problem of the mass transfer can be decomposed to the following
parts: the mass transfer from two circumferences and the mass transfer from the film. We will
examine
these problems with the help of the lattice Boltzmann method and compare them against analytical
solutions. The next sections will present the lattice Boltzmann method and a few benchmarks.

@@

The TRT collision operator \cite{ginzburg-boundary-main}
decomposes the populations and the equilibrium
distribution into the symmetric and antisymmetric parts:
\begin{equation}
\label{trtdecomp}
f^{\pm}_i=\frac{f_i\pm f_{\bar{i}}}{2}\;,\; 
{eq_i}^{\pm}=\frac{eq_i\pm eq_{\bar{i}}}{2}\;,
\end{equation}
where $\bar{i}$ is the opposite direction to the $i$-th direction.
The collision is performed with two independent relaxation rates for 
symmetric and antisymmetric modes:
\begin{equation}
\label{trt}
\begin{aligned}
&f_i^{*}(\bm{x},t)=f_i(\bm{x},t)-\omegaplus (f_i^{+} - eq_i^+)-\omegaminus
(f_i^{-} -
eq_i^-)\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t).
\end{aligned}
\end{equation}
Note that the BGK collision operator is the particular subclass of the TRT relaxation operator with
$\omegaplus=\omegaminus$. In comparison with the BGK collision operator,
the TRT collision operator has the additional degree of freedom. Thus, the TRT operator then
introduces
the free parameter
$\Lambda=\Bigl(\frac{1}{\omegaplus}-\frac{1}{2}\Bigr)\Bigl(\frac{1}{\omegaminus}-\frac{1}{2}
\Bigr)$. 
This free parameter controls the effective location of the bounce-back
walls \cite{ginzburg-multireflection}, second-order accuracy of
boundary \cite{ginzburg-boundary-main} and interface schemes \cite{ginzburg-discontinious}, 
spatial accuracy \cite{ginzburg-recurrence,servan-trt-stability},
consistency \cite{ginzburg-brinkman} and, to some extent,
stability \cite{kuzmin-stability-optimal,kuzmin-d1q3,servan-trt-stability}.
In particular, $\Lambda=\frac{1}{4}$ achieves the optimal stability for the
linear advection-diffusion isotropic equation \cite{kuzmin-stability-optimal}.

@@

The equilibrium functions for $D2Q9$ TRT model are represented as \cite{kuzmin-stability-optimal}:
\begin{equation}
\begin{aligned}
&eq_i^{+}=eq_i^{(m)}+g^{(u)} eq_i^{(u)}\\
&eq_i^{(m)}=t_i^{(m)} c_e+ eq_i^{(a)}\\
&eq_i^{(u)}=t_i^{(u)} \frac{u_x^2+u_y^2}{2}+\frac{u_x^2-u_y^2}{4} p_i^{(xx)}+g_{xy}^{(u)}\frac{u_x
u_y}{4} p_i^{xy}\\
&eq_i^{(a)}=\frac{K_{xx}-K_{yy}}{4} p_i^{xx}+\frac{D_{xy}}{4} p_i^{(xy)},
\end{aligned}
\end{equation}
where $K_{xx,yy,xy}$ are proportional to components of the diffusion tensor,
$c_e=\frac{K_{xx}+K_{yy}}{2}$, parameters $g^{(u)}$ and $g^{u}_{xy}$ are either zero or one (see
below), the tensor $p_i^{(xx)}=c_{ix}^2-c_{iy}^2$, the tensor $p_i^{(xy)}=c_{ix} c_{iy}$, the
weights
$t_i^{(u,m,a)}$ can be chosen based on stability criteria. Though not the best set of weights for
stability \cite{kuzmin-stability-optimal}, but the most spread set of weights, so called
``hydrodynamic`` weights, was chosen:
\begin{equation}
t_i^{(u)}=t_i^{(m)}=t_i^{(a)}=\Bigl\{0,\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{12},
\frac {1}{12},\frac{1}{12},\frac{1}{12}\Bigr\}
\end{equation}

@@

Boundary conditions for the mass transfer used in this work are all of Dirichlet type. We validated
two types of boundary conditions: Inamuro boundary conditions \cite{inamuro-scalar-boundary} and
pressure anti bounce-back boundary conditions \cite{ginzburg-multireflection}. However, the
simulation results presented in this work are only pressure anti bounce-back due to its simplicity
to handle any complicated type of boundary:
\beq
\label{antibb}
f^{*}_{B,i}=-f^{*}_{F,\bar{i}} + 2 eq^+(\cstar,\bm{u}),
\feq
where $\cstar$ is the concentration to be imposed at the surface, $\bm{u}$ is the surface velocity,
$i$ is the direction number pointing to the domain (unknown) located at the boundary surface $B$,
$\bar{i}$ is the direction number opposite to $i$ and is located at the fluid $F$ specifically that
$B$ is located at the location $F+c_i$.

Note that the parameters of the lattice
Boltzmann scheme are connected with the physical parameters only through the non-dimensional
numbers governing the physics of the problem. In our case, this number is the Peclet number, $Pe=\frac{\ububble L}{D}$.
Therefore, one can substitute any quantity, i.e.
$\ububble$ in the lattice Boltzmann units as soon as the
Peclet number is the matched in physical space and numerical simulations. This fact that $\ububble$
can be chosen arbitrarily is extremely useful in the context of numerical simulations, since it can
increase the time step, or decrease the computational demand by order of magnitude to obtain
meaningful results. This point will be extensively covered later.

Next section will cover LBM benchmarks that resemble the mass transfer problem from a bubble, as as
mass transfer from the film  and circumferences.

>>>>

\begin{document}
\maketitle
\begin{abstract}
This work presents a thorough procedure of how to determine the volumetric mass transfer
coefficient
in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train flow. In this work we address the situation when the hydrodynamic pattern changes from having a vortex in slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}. For the latter case the bubble shape is not symmetric and cannot be approximated through flat surfaces and circumferences as it is done in literature \cite{vanbaten-circular,kreutzer-overview}. For $Ca<0.7$ where there is a vortex in the slug the tracer is well mixed. Thus, it is common to  use periodic boundary conditions  and the inlet/outlet averaged concentration as the characteristic concentration, Eq. \ref{eq:main:definition}.  The later is not valid for flows where tracer is not well mixed, i.e. $Ca>0.7$. This work through examination of different boundary conditions and different characteristic concentration definitions establishes the procedure to determine the volumetric mass transfer coefficient for flows with the capillary number $0.1<Ca<1.0$. With simulations of a few unit cells the connection between time (used in simulations) and space (used in experiments) domains is established. The following boundary conditions were examined: periodic boundary conditions, open boundaries, and a few unit cells open boundary
simulations. It was shown that the time dependent average concentration to be a characteristic
concentration produce robust results, though a bit overpredicted. The inlet/outlet flux averaged
concentration \cite{vanbaten-circular} to be a characteristic concentration produce unreliable results in the range of capillary numbers specified. In comparison with works \cite{vanbaten-circular,kreutzer-overview} where only one unit cell was simulated, a few
unit cells open boundaries conditions resemble physical situation most closely. However, they  are harder to be conducted 
but provide the volumetric mass transfer coefficient based only on the spatial information for the
capillary number $Ca>0.7$. 
%As well, a few unit cells simulations can provide more information about
%mixing in the liquid slug in comparison with the periodic boundary conditions. 
We show
that all presented in literature strategies are extreme limits of one equation,
Eq. \ref{main:main:main}. Finally, simulation results for different Peclet numbers
are compared with analytical predictions presented by \citet{vanbaten-circular} and shows a good
agreement. The work is of use to people performing mass transfer numerical simulations for bubble
train flow in microchannels and/or within the
lattice Boltzmann model context.
\end{abstract}

\section{Introduction}
\label{intro}
The monolith reactors during last years are getting more attention as a promising alternative to slurry
reactors and trickle bed reactors \cite{kreutzer-overview,bercic-mass}. Reactors usually operate in the Bretherton-Taylor regime \cite{bretherton,taylor} which describes a flow through
a liquid medium of equally sized, long air bubbles, see
Fig. \ref{fig:benchmark:hydro}. This flow regime
is characterized by surface tension dominance over inertia and viscous effects, and by
comparatively small gas flow velocities \cite{yue-mass}. Due to surface tension dominance bubble train flow exhibit advantageous properties which can not be achieved
in macroscopic counterparts: liquid thin films \cite{bretherton} between bubbles and walls remarkably enhance mass
transfer from a gas and walls to liquid; the plug flow regime occurring in monolith reactors allows to perform chemical reactions in slugs only \cite{kreutzer-overview}. Moreover, the low slip velocity between gas and liquid is utilized in the experiments to measure liquid velocity \cite{taylor}:  bubbles travelling with approximately the same velocity as liquid can be captured with a camera. Overall, nowadays one can find a large number of applications
of the Bretherton-Taylor bubble train flow: continuous flow analyzers to measure liquid velocity, lung openings, chemical reactions
as hydrogenation of nitroaromatics, 2-ethyl-hexenal, Fischer-Tropsch synthesis, etc. Extensive reviews of \citet{kreutzer-overview,
gupta-review,yue-mass} cover a significant number of applications.

This work is focused on the gas to liquid mass transfer 
study for the Bretherton/Taylor flow. The good understanding of mass transfer and how it depends on the
parameters as the capillary number, the Reynolds number, slug and bubble lengths allows to properly
manufacture a microchannel with necessary properties to ensure that chemical
reactions are performed in the best manner. The mass transfer coefficient characterizing mass transfer is defined as the flux from the surface divided on the concentration difference between the concentration imposed on the body and the characteristic concentration in the domain. The characteristic concentration in the domain significantly depends on underlying hydrodynamics fields.  For example, experimental studies \cite{yue-mass,bercic-mass} show a complex dependency of the mass transfer coefficient on flow parameters: bubble and slug lengths, bubble velocity, which are in turn depend on the capillary number $Ca$ and the Reynolds number $Re$. 
\citet{yue-mass} established an experimental correlation for the volumetric mass transfer coefficient for a bubble train as
a function of the diffusion coefficient, slug and bubble lengths, and bubble velocity: 
\begin{equation}
\vol =\frac{2}{d_h} \Bigl(\frac{D
\ububble}{\lbubble+\lslug}\Bigr)^{0.5}
\Bigl(\frac{\lbubble}{\lbubble+\lslug}\Bigr)^{0.3},
\end{equation}
where $\vol$ is the volumetric mass transfer coefficient, $d_h$ is the hydrodynamic capillary
diameter, $\lbubble$ is the bubble length, $\lslug$ is the slug distance (between bubbles),
$\ububble$ is the bubble velocity, $D$ is the diffusion coefficient. 
%This correlation is within $10\%$ accuracy for experimental
%results for the microchannels with rectangular/square crosssections for the following range of
%parameters: bubble velocity $0.4\,\mathrm{m/s}<\ububble<2\,\mathrm{m/s}$, bubble length
%$1.4<\frac{\lbubble}{d_h}<6.3$, slug length $1<\frac{\lslug}{d_h}<3.2$,
%hydrodynamic diameter $d_h = 400 \mathrm{\mu m}$.

Thus, the understanding of mass transfer for the bubble train flow is not possible without deep understanding of hydrodynamic patterns. There are a number of studies available for the hydrodynamic study of the bubble train flow: experimental \cite{kreutzer-pressure-drop,cerro-space,cerro-bubble-train} and numerical \cite{wang-non-circular,kuzmin-binary3d,giavedoni-numerical,heil-threedim}. For the flow of long bubbles between parallel plates chosen here as the study case it is indicated that  there exists a vortex in the liquid slug for $Ca<0.7$ , see Fig. \ref{fig:streamlines:tweaked:velocity}.  As well, the bubble shape for low capillary numbers ($Ca<0.1$ \cite{cerro-bubble-train}) is symmetric. Thus, in the case of flow between plates for $Ca<0.1\div0.7$ the bubble shape can be represented as two hemicircles and two planar interfaces with the vortex existing in the liquid slug. This is heavily utilized for mass transfer analytical estimations.

As mentioned  the mass transfer coefficient is defined as the mass flux from certain area, Eq. \ref{eq:main:definition}. Thus, analytical estimations \cite{kreutzer-overview,irandoust} are  based on decomposition of the bubble shape. The mass transfer coefficient is calculated through two separate contributions from two films and two hemicircles. For both contributions the Higbie penetration theory \cite{higbie} is utilized, which states that the mass transfer coefficient  for simple geometry flow depends on the average time of liquid packet interaction with a geometrical feature. It can be calculated as $\sqrt{\frac{\pi D}{t_{\mathrm{char}}}}$, where $t_{\mathrm{char}}$ is the interaction characteristic time, $D$ is the diffusion coefficient. As the example for the Higbie penetration theory, the mass  transfer coefficient for the flow of bubbles between parallel plates  is calculated as (similarly to work of \citet{vanbaten-circular}):
\beq
k_L=2 \sqrt{\frac{\pi D}{t_{\mathrm{film}}}}+2 \sqrt{\frac{\pi D}{t_{\mathrm{circle}}}},
\feq
where $t_{\mathrm{film}}=\frac{\lfilm}{\ububble}$ states for the interaction time of liquid traveling through planar part of the bubble, $t_{\mathrm{circle}}=\frac{\pi R_{\mathrm{circle}}}{\ububble}$ is the time which liquid in the slug travels the distance of half bubble circumference.

@@

However, available numerical studies of mass transfer \cite{kreutzer-overview,vanbaten-circular} are lacking the simulation of  bubble shapes. The usual simulation of the mass transfer  is performed as follows: 
\begin{description}
\item[I] The bubble shape is calculated either through analytical correlations \cite{bretherton} or experimental correlations \cite{cerro-bubble-train} without directly resolving bubbles shapes through multiphase simulations.  The expressions for bubble shapes are available only for flows with the capillary number $Ca<0.1$.
\item[II] Hydrodynamics fields are then obtained by performing simulations of one-component flow around the droplet by imposing the bubble velocity on the walls. Thus, the simulations are performed in the reference frame moving with the bubble. The stress free condition is imposed at the bubble surface. 
\item[III]  The mass transfer simulations are  performed in the reference frame moving with the bubble. The saturation concentration is imposed at the bubble surface. Only one unit cell with one bubble in it is used for simulations. Periodic boundary conditions are utilized to determine the volumetric mass transfer coefficient, which is calculated through the following equation \cite{vanbaten-circular}:
\label{main:simulation:equation}
\vol=\frac{\mathrm{\overline{Flux}}}{\cbubble-\langle\coutlet\rangle} \frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\
cell\ volume}},
where $\langle\coutlet(t)\rangle=\int{C \uoutlet \mathrm{d}A}/\int{\uoutlet\mathrm{d}A}$ is the averaged in space outlet concentration as the function of time used as the characteristic concentration in the definition, Eq. \ref{eq:main:definition}. 
The averaged in time concentration flux ($\mathrm{\overline{Flux}}$) is calculated as the difference between the overall
average concentration in the whole domain ($\langle\coverall\rangle=\int_{V} C \mathrm{d}V /V$)
at time
$t_1$ and at time $t_2$ divided on the time difference $t_2-t_1$. The agreement between numerical
simulations \cite{vanbaten-circular} and correlations of \citet{bercic-mass} was good. 
\end{description}
There is a certain critics towards the presented numerical approaches \cite{vanbaten-circular,kreutzer-overview}. They mainly originate from the bubble shape approximation. It is taken symmetrical, i.e. consisting from hemispheres and cylinder film for the case of flow in circular capillaries. This is valid  for small capillary numbers $Ca<0.1$. As it was discussed, for such capillary numbers in the reference frame moving with the bubble there is a vortex in slug. Thus, the tracer is well mixed in slug.The choice of the characteristic concentration needed for the mass transfer coefficient, Eq. \ref{eq:main:definition}, in this case  is obvious. With small differences in results it can be either averaged in the liquid slug concentration or the outlet space averaged concentration as used in formulation of \citet{vanbaten-circular}. Another critics is towards periodic boundary conditions to calculate the mass transfer coefficient. While it is clear to use periodic boundary conditions for the calculation of hydrodynamics fields for  long bubbles motion in the microchannel, it is not the case for mass transfer simulations. Experimental correlations \cite{bercic-mass} show that the concentration along the streamwise direction changes as the exponential function. However, mass transfer simulations are made only for one unit cell using the periodic boundary conditions with the same concentration at the inlet and at the outlet. The question how  one unit cell simulation corresponds to experimental measurements arises where concentration difference is measured at the distances of at least of a few unit cells difference \cite{bercic-mass}. In other words one needs to understand how the discrete one unit cell simulation corresponds to the continuous picture in experiments where one does not distinguish bubbles but takes measurements of concentration at different locations.

@@

\item
The approach of \citet{vanbaten-circular} is examined. Periodic boundary conditions were used. The work \cite{vanbaten-circular}
calculates the mass transfer coefficient as the gain of the mass in the system
divided by the concentration difference
multiplied on the surface area:
\beq
\label{eq:vanbaten:formulation}
\vol=\frac{\dot{m}}{P \Delta C}\frac{P}{V}=\frac{\dot{m}}{V (\cstar-\langle C(t) \rangle)},
\feq
where mass difference in the domain can be calculated as:
\beq
\dot{m}=\frac{m_2-m_1}{t_2-t_1}=\frac{\int_{liq}{C(\bm{x},t_2)\mathrm{d}V}-\int_{liq}{
C(\bm{x},t_1)\mathrm { d } \bm{x}} } {
t_2-t_1 }.
\feq
In the approach of \citeauthor{vanbaten-circular} the inlet and outlet concentrations were taken as the
characteristic concentration $\langle C(t) \rangle $.

@@

\section{Validation}
As it was discussed analytical correlations for the mass transfer coefficient are derived  as two separate contributions: the mass transfer from two circumferences and the mass transfer from the film. We will
examine
these problems closely with the help of the lattice Boltzmann method and compare them against analytical
solutions. The next sections will give a short introduction on the lattice Boltzmann method and present a few benchmarks.

@@

The TRT collision operator \cite{ginzburg-boundary-main}
decomposes the populations and the equilibrium
distribution into the symmetric and antisymmetric parts:
\begin{equation}
\label{trtdecomp}
f^{\pm}_i=\frac{f_i\pm f_{\bar{i}}}{2}\;,\; 
{eq_i}^{\pm}=\frac{eq_i\pm eq_{\bar{i}}}{2}\;,
\end{equation}
where $\bar{i}$ is the opposite direction to the $i$-th direction.
The collision is performed with two independent relaxation rates for 
symmetric and antisymmetric modes:
\begin{equation}
\label{trt}
\begin{aligned}
&f_i^{*}(\bm{x},t)=f_i(\bm{x},t)-\omegaplus (f_i^{+} - eq_i^+)-\omegaminus
(f_i^{-} -
eq_i^-)\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t).
\end{aligned}
\end{equation}
Note that the BGK collision operator is the particular subclass of the TRT relaxation operator with
$\omegaplus=\omegaminus$. In comparison with the BGK collision operator,
the TRT collision operator has the additional degree of freedom. Thus, the TRT operator 
introduces
the free parameter
$\Lambda=\Bigl(\frac{1}{\omegaplus}-\frac{1}{2}\Bigr)\Bigl(\frac{1}{\omegaminus}-\frac{1}{2}
\Bigr)$. 
This free parameter controls the effective location of the bounce-back
walls \cite{ginzburg-multireflection}, second-order accuracy of
boundary \cite{ginzburg-boundary-main} and interface schemes \cite{ginzburg-discontinious}, 
spatial accuracy \cite{ginzburg-recurrence,servan-trt-stability},
consistency \cite{ginzburg-brinkman} and, to some extent,
stability \cite{kuzmin-stability-optimal,kuzmin-d1q3,servan-trt-stability}.
In particular, $\Lambda=\frac{1}{4}$ achieves the optimal stability for the
linear advection-diffusion isotropic equation \cite{kuzmin-stability-optimal}.

@@

The equilibrium functions for $D2Q9$ TRT model are represented as \cite{kuzmin-stability-optimal}:
\begin{equation}
\begin{aligned}
&eq_i^{+}=eq_i^{(m)}+g^{(u)} eq_i^{(u)}\\
&eq_i^{(m)}=t_i^{(m)} c_e+ eq_i^{(a)}\\
&eq_i^{(u)}=t_i^{(u)} \frac{u_x^2+u_y^2}{2}+\frac{u_x^2-u_y^2}{4} p_i^{(xx)}+g_{xy}^{(u)}\frac{u_x
u_y}{4} p_i^{xy}\\
&eq_i^{(a)}=\frac{K_{xx}-K_{yy}}{4} p_i^{xx}+\frac{K_{xy}}{4} p_i^{(xy)},
\end{aligned}
\end{equation}
where $K_{xx,yy,xy}$ are proportional to components of the diffusion tensor,
$c_e=\frac{K_{xx}+K_{yy}}{2}$, parameters $g^{(u)}$ and $g^{u}_{xy}$ are either zero or one (see
below), the tensor $p_i^{(xx)}=c_{ix}^2-c_{iy}^2$, the tensor $p_i^{(xy)}=c_{ix} c_{iy}$, the
weights
$t_i^{(u,m,a)}$ can be chosen based on stability criteria. The most spread set of weights, so called
``hydrodynamic`` weights, was chosen:
\begin{equation}
t_i^{(u)}=t_i^{(m)}=t_i^{(a)}=\Bigl\{0,\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{3},\frac{1}{12},
\frac {1}{12},\frac{1}{12},\frac{1}{12}\Bigr\}
\end{equation}

@@

Boundary conditions for the mass transfer used in this work are all of Dirichlet type. We validated
two types of boundary conditions: Inamuro boundary conditions \cite{inamuro-scalar-boundary} and
pressure anti bounce-back boundary conditions \cite{ginzburg-multireflection}. However, the
simulation results presented in this work are only pressure anti bounce-back due to its simplicity
to handle any complicated type of boundary:
\beq
\label{antibb}
f^{*}_{B,i}=-f^{*}_{F,\bar{i}} + 2 eq^+(\cstar,\bm{u}),
\feq
where $\cstar$ is the concentration to be imposed at the surface, $\bm{u}$ is the surface velocity,
$i$ is the direction number pointing to the domain (unknown) located at the boundary surface $B$,
$\bar{i}$ is the direction number opposite to $i$ and is located at the fluid $F$ specifically that
$B$ is located at the location $F+\bm{c_i}$.

Note that the parameters of the lattice
Boltzmann scheme are connected with  physical parameters only through  non-dimensional
numbers governing the physics of the problem. In our case, this number is the Peclet number, $Pe=\frac{\ububble L}{D}$.
Therefore, one can substitute any quantity, i.e.
$\ububble$ in the lattice Boltzmann units as soon as the
Peclet number is  matched in physical space and numerical simulations. This fact that $\ububble$
can be chosen arbitrarily is extremely useful in the context of numerical simulations, since it can
increase the time step, or decrease the computational demand by order of magnitude to obtain
meaningful results. This point will be used in simulations and covered later.

Next section will cover LBM benchmarks that resemble the mass transfer problem from a bubble, as 
mass transfer from the film  and circumferences.
